---
title: "STAT 440 Statistical Data Management - Fall 2021"
output: html_document
---

## Homework05
### Due: Monday November 15, 2021 11:59 pm US Central Time
#### Created by Course Staff

Grading Rubric (per question):  
2 points if complete and correct  
1 point if incomplete or incorrect  
0 points if no attempt made  


**Retrieving your work**

This and all future homework .md files are written in Markdown. The .md file is rendered as an .html file for easier viewing. This and all future homework assignments are located in the **homework** directory within the **stat440-fa21-course-content** repo, i.e. **stat440-fa21-course-content/homework** in GHE. It is always recommended that you **pull** (or refresh) the **stat440-fa21-course-content** repo to ensure that you have the most updated version of all course content including the homework assignments. After pulling (or refreshing) the **stat440-fa21-course-content** repo, the homework file will be in the homework directory. Once you have accessed the .md file, copy all text from the .md file and paste all text into your reproducible document file (either .Rmd or .ipynb file). 

**Submitting your work**

In your individual student repo (named as your netID), you are to submit ***two*** files:

a. Your reproducible document file (either .Rmd or .ipynb file) which should be saved as homework##-netID.Rmd or homework##-netID.ipynb depending on your preferred software. For example, my homework 01 file would be saved as homework01-kinson2.Rmd.

b. Your rendered reproducible document file (.html) which should be saved as homework##-netID.html. For example, my homework 01 file would be saved as homework01-kinson2.html.

**Do not place your files in a sub-directory (or folder).** You have an unlimited number of submissions, but only the latest proper submission (commit and push) will be viewed and graded. You are allowed to only use one software language per homework file. This means that you cannot mix software languages (R and Python) in one single homework file. **Remember the .Rmd (or .ipynb) file needs to render properly to .html before submitting.** 


***

Undergraduates are expected to complete the problems **#1-#8**. Graduate students are expected to complete all problems.

**The following problems should be completed by you as an individual. If any problem asks you a particular question, be sure to answer it completely (with code, written sentences, or both). Written sentences should not appear in code chunks or code cells. Written sentences should appear in Markdown syntax unless specifically stated otherwise.**

***Do not change anything in this file above the double line. Doing so will prevent your homework from being graded.***

***
***

### Use only one programming language for this entire homework assignment. Use the URLs to access the data (if any). No local files allowed.

**#1** (Markdown) Using Markdown syntax, make a non-numbered list with your first name in normal text as the first item and your last name in bold text as the second item.

- Kamila
- **Makhambetova**



**#2** (Personal survey, Markdown) Answer the following survey questions in your honest opinion using Markdown syntax and complete sentences.

- Is this course too easy, easy, not easy and not hard, hard, or too hard? 
<br>
This course is not easy and not hard, but requires to learn material that was not covered in the class.

- Has the difficulty level increased since the beginning of the semester?
<br>
The difficulty level increased a lot since the beginning of the semester.

- Do you feel as though you can start working as a data engineer or data wrangler after taking this course because of the course's content?
<br>
I am not sure as I can start working as a data engineer or data wrangler after taking this course because of the course's content. But hope this course knowledge will be useful during working as data scientist. 


**#3** (Validating data, Cleaning data, Markdown) 

statement: The **WHERE** clause allows conditions to be placed on the groups for group processing. 

If the text in bold is the term that makes the statement true, then write TRUE below in all caps. If the text in bold is the term that makes the statement false, then write FALSE below in all caps and write the correct term in bold text in a new line under FALSE.
<br>
<br>
FALSE
<br>
**HAVING**
<br>
<br>


**#4** (SQL Subqueries, Markdown) 

statement: An SQL subquery can take place within the **FROM** clause. 

If the text in bold is the term that makes the statement true, then write TRUE below in all caps. If the text in bold is the term that makes the statement false, then write FALSE below in all caps and write the correct term in bold text in a new line under FALSE.
<br>
<br>
TRUE
<br>
<br>


**#5** (Accessing and importing data) Import the following dataset using a programming language and their corresponding data URLs, and print the structure of each. Read the data description to help verify the import was successful. 

- Champaign County Sheriff Office & Jail Data. [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/blob/master/data/ccso-jail-data.dat) or [Box data URL](https://uofi.box.com/shared/static/lurfpdqk63au11yzra4a88sxizda0iz5.dat).

  - The dataset (a .dat file) contains over 11082 observations and 40 columns for individuals who were booked into jail in Champaign County. The individuals are not identifiable based on personal identity, but they are given unique jacket numbers. Some rows in the data appear more than once because of multiple crimes being assigned to one person. But there is also the possibility of there being recidivists. A recidivist is a person who goes to jail repeatedly (more than one datetime). The demographics of the people, reasons for being booked, and crime code are also given in the data. The original source is the Champaign County Sheriff Office & Jail (CCSO).

- 2020 NBA Season Data (existing as four separate data sets). [stint1 - GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/nba2020-2021-stint01-data.json), [stint1 - Box data URL](https://uofi.box.com/shared/static/fk1q6htmes532jz2obi89zj2v7bobh37.json), [stint2 - GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/nba2020-2021-stint02-data.json), [stint2 - Box data URL](https://uofi.box.com/shared/static/aix57cgzs4bhzsvorpv924rrv1w57wod.json), [stint3 - GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/nba2020-2021-stint03-data.json), [stint3 - Box data URL](https://uofi.box.com/shared/static/6a4nbioxh8yek5oppb06jn4429n7fekr.json), [stint4 - GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/nba2020-2021-stint04-data.json), and [stint4 - Box data URL](https://uofi.box.com/shared/static/9ap7efzs42xldrhr1cpt1ql53ip1v25u.json).

  - The dataset (four .json files) contains a total of 540 unique observations and 25-26 columns. The observations are professional basketball players in the National Basketball Association (NBA) and their season statistics for the 2020-2021 season. A stint is the duration of time someone worked for a particular employer. In this basketball data, a stint is a number representing the time a player was with a particular team. Some players are traded during the season, which means they would have more than 1 stint. A player's basketball statistics start over for each stint they have in a season. There are 540 unique player IDs across all stints. The original source is the NBA and Sports Reference.
  
  - [NBA Data key - GHE link](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/nba2020-2021-data-key.pdf) or [NBA Data key - Box link](https://uofi.box.com/shared/static/vcz69vd5v51aokatt6bafmzdge3jhwi0.pdf)

```{r}
library(tidyverse)
library(jsonlite)
jail<-read_csv("https://uofi.box.com/shared/static/lurfpdqk63au11yzra4a88sxizda0iz5.dat")
NBA1<-fromJSON("https://uofi.box.com/shared/static/fk1q6htmes532jz2obi89zj2v7bobh37.json")
NBA2<-fromJSON("https://uofi.box.com/shared/static/aix57cgzs4bhzsvorpv924rrv1w57wod.json")
NBA3<-fromJSON("https://uofi.box.com/shared/static/6a4nbioxh8yek5oppb06jn4429n7fekr.json")
NBA4<-fromJSON("https://uofi.box.com/shared/static/9ap7efzs42xldrhr1cpt1ql53ip1v25u.json")
str(jail)
str(NBA1)
str(NBA2)
str(NBA3)
str(NBA4)

```



**#6** (SQL Weeks 03-06, SQL Weeks 09-10, SQL Subqueries) Champaign, Danville, Rantoul, Savoy, and Urbana are just five towns in Champaign County. Using the CCSO Jail Data imported in **Problem 5** and a programming language applying data validating and cleaning and other data management concepts, create a new data object called `cc` that contains any person booked into jail who resides in these five towns. Afterwards, arrange by the city name in alphabetical order and select the following columns: jacket number, age at arrest, city, state, sex, race. Now, print the first 10 observations of `cc`. **In this problem, you should make sure the five towns are correctly spelled. If they are not correctly spelled, then your code should correct the misspelled towns. You do not have to remove duplicates in this problem. You must use SQL to complete this problem.**

```{r}
library(sqldf)
library(tidyverse)



cc<-sqldf("SELECT * 
  FROM (SELECT jacket_number,`age arrest`,state,sex,race,
  CASE 
   WHEN city = 'CHAAMPAIGN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMAIGN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMAPAIGN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMAPIAGN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMAPIGN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMPAGIN' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMPAING' THEN 'CHAMPAIGN'
   WHEN city = 'CHAMPIAGN' THEN 'CHAMPAIGN'
   WHEN city = 'DANVILLIE' THEN 'DANVILLE'
   WHEN city = 'DAVILLE' THEN 'DANVILLE'
   WHEN city = 'RANOUL' THEN 'RANTOUL'
   WHEN city = 'RANTOU' THEN 'RANTOUL'
   WHEN city = 'SAVOYU' THEN 'SAVOY'
   WHEN city = 'URANA' THEN 'URBANA'
   WHEN city = 'URBNANA' THEN 'URBANA'
   ELSE city
   END AS city
FROM jail)
  WHERE city = 'CHAMPAIGN' OR city= 'DANVILLE' OR city= 'RANTOUL' OR city= 'SAVOY'
  OR city= 'URBANA'
  ORDER BY city
")

head(cc,10)


```




**#7** (SQL Weeks 03-06, SQL Weeks 09-10, SQL Subqueries, Markdown) For our purposes, let's define recidivism as the act of a single person going to jail more than one time. A recidivist is a person who goes to jail repeatedly (more than one datetime). Use the CCSO Jail Data imported in **Problem 5**, and a programming language applying data management concepts to answer the following question in Markdown syntax and complete sentence(s): 

- How many people booked into jail are recidivists?

**You must answer in Markdown syntax in a complete sentence and show code as evidence of your answer. You must use SQL to complete this problem.**
<br>
<br>
**Answer**
<br>
There are 909 recidivists who were booked into jail.

```{r}


recidivists1<-sqldf(" SELECT jacket_number, COUNT(jacket_number) AS times_in_jail
      FROM (SELECT jacket_number, booking_date
      FROM jail
      GROUP BY jacket_number, booking_date)
      GROUP BY jacket_number
      HAVING times_in_jail>1
      ")

nrow(recidivists1)


```



**#8** (Regular Expressions and String Manipulation) Beginning with the NBA data sets in **Problem 5** and the information in the data key in **Problem 5**, use a programming language to split the Player column of the fourth NBA dataset into two new columns: PlayerID and PlayerName such that the original Player column is removed and these two new columns are the first two columns of the fourth NBA data. **These two columns should look similar to the first two columns of the first NBA dataset.** Now, print the first 10 rows of the fourth and first NBA datasets, respectively. **SQL is not required for this problem.**

```{r}
NBA4<-fromJSON("https://uofi.box.com/shared/static/9ap7efzs42xldrhr1cpt1ql53ip1v25u.json")
Coordinates<-str_split(NBA4$Player, "\\\\", simplify = TRUE)
#head(Coordinates, n=10)

PlayerName1<-Coordinates[ ,1]
PlayerID1<-Coordinates[ ,2]

NBA4<-NBA4 %>%select(-Player)

NBA4_new <- NBA4 %>%
  mutate(PlayerName=PlayerName1, PlayerID=PlayerID1) %>%
  select(PlayerID, PlayerName, colnames(NBA4))

head(NBA4_new, 10)
head(NBA1,10)




```



**#9** (SQL Weeks 03-06, SQL Weeks 09-10, SQL Subqueries) Beginning with the three NBA data sets in **Problem 5**, the newly created fourth NBA dataset in **Problem 8**, and the information in the data key in **Problem 5**, do all of the following:

- combine the four data sets

- summarize the data such that all basketball statistics are totaled across all stints per player; there should only be 540 players after summarizing; use the final stint's Position, Age, and Team for each player

- mutate a new column called "REBOUNDS" that represents a combination of offensive rebounds and defensive rebounds

- mutate a new column called "EFFICIENCY" that represents the NBA's efficiency rating using the formula: (points + rebounds + assists + steals + blocks - ((field goals attempted - field goals made) + (free throws attempted - free throws made) + turnovers)) / games played

- mutate a new column called "ATO" that represents the assists to turnovers ratio (assists divided by turnovers)

- mutate a new column called "PTM" that represents the points to minutes played ratio (points divided by minutes played)

- mutate a new column called "3PP" that represents three-point shots made out of three-point attempts multiplied by 100

- de-select the tm and stint columns

Now, print the first 15 observations of the resulting single dataset such that the data is sorted by Rebounds in descending order then Player ID alphabetically and showing the following columns only: PlayerName, PlayerID, Pos, Age, Rebounds. **You must use SQL to solve this problem.**

```{r}


NBA_total<-sqldf(" SELECT *
                      FROM NBA1
                      UNION ALL
                      SELECT *
                      FROM NBA2
                       UNION ALL
                      SELECT *
                      FROM NBA3
                      UNION ALL
                      SELECT *
                      FROM NBA4_new")



NBA_total2<-sqldf(" SELECT PlayerID, PlayerName, Pos, Age, Tm, SUM(G) AS G, SUM(GS) AS GS, SUM(MP) AS MP, SUM(FG) AS FG, SUM(FGA) AS FGA, SUM([3P]) AS [3P], SUM([3PA]) AS [3PA], SUM([2P]) AS [2P], SUM([2PA]) AS [2PA], SUM(FT) AS FT, SUM(FTA) AS FTA, SUM(ORB) AS ORB, SUM(DRB) AS DRB, SUM(TRB) AS TRB, SUM(AST) AS AST, SUM(STL) AS STL, SUM(BLK) AS BLK, SUM(TOV) AS TOV, SUM(PF) AS PF, SUM(PTS) AS PTS, SUM(stint) AS stint
FROM NBA_total
GROUP BY(PlayerID)
HAVING MAX(stint)
")

NBA_total3<-sqldf(" SELECT *, ORB+DRB AS REBOUNDS, (PTS+ORB+DRB+AST+STL+BLK-(FGA-FG)+(FTA-FT)+TOV)/G AS EFFICIENCY, AST/TOV AS ATO, PTS/MP AS PTM, ([3P] * 100)/[3PA] AS '3PP' FROM NBA_total2
                  ")

NBA_total3<-sqldf(" SELECT PlayerID, PlayerName, Pos, Age, G, GS, MP, FG, FGA, '3P', '3PA', '2P', '2PA', FT, FTA, ORB, DRB, TRB, AST, STL, BLK, TOV, PF, PTS, REBOUNDS, EFFICIENCY, ATO, PTM, '3PP'
FROM NBA_total3")


sqldf(" SELECT PlayerName, PlayerID, Pos, Age, REBOUNDS
FROM NBA_total3
ORDER BY REBOUNDS DESC, PlayerID
LIMIT 15")



```



**#10** (SQL Weeks 03-06, Markdown) Using the data in **Problem 9**, the information in the link in **Problem 5**, and a programming language applying data management concepts with SQL to answer the following question in Markdown syntax and complete sentence(s): 

- Which (if any) of the 15 most efficient players made the 2020 US Olympic Basketball Team?

**You must answer in Markdown syntax in a complete sentence and show SQL code as evidence of your answer. You must use SQL to solve this problem. This link https://www.teamusa.org/tokyo-2020-olympic-games/meet-team-usa/athletes may be of assistance in identifying 2020 US Olympic Basketball Team.**
<br>
<br>
**Answer**
<br>
None of 15 most efficient players were part of the 2020 US Olympic Basketball Team.

```{r}

effic<- sqldf(" SELECT *
FROM NBA_total3
ORDER BY EFFICIENCY
LIMIT 15")

Bask_team<-sqldf(" SELECT *
FROM effic
WHERE PlayerName IN ('Bam Adebayo', 'Devin Booker', 'Kevin Durant', 'Jerami Grant', 'Draymond Green', 'Jrue Holiday', 'Keldon Johnson', 'Zach LaVine', 'Damian Lillard', 'JaVale McGee', 'Khris Middleton', 'Jayson Tatum')
")

nrow(Bask_team)

```


