---
title: "Lab 2"
author: "Kamila Makhambetova"
date: "01 09 2021"
output: html_document
---

# Problem 1

List of names: 

**1. Kamila Makhambetova** 

2. Harry  Liang 

3. Zhen Zhou 

4. Jasmine Zhang

5. Neha Bharadwaj   

6. Jefferson Mathews 

7. Niko Yin 

8. Yanyi Lu 

# Pronlem 2
<br>
Each lab assignment worths 3.3 points, as each person needs to submit 3 lab assignments and total weight of lab assignments is 10.
<br>

# Problem 3
<br>
A toenized URL is c.
<br>

# Problem 4
List of questions:

1.  How many lines does scarf have?

2. What is the thing in the middle of piccture?

3. What is the category of the data?

# Problem 5
```{r}
library(readr)
rows <- read_csv("https://data.urbanaillinois.us/api/views/6gtk-bwms/rows.csv?accessType=DOWNLOAD")
print(rows)

```

# Problem 6


```{r}

raw2<- read_tsv("https://data.urbanaillinois.us/api/views/tsn9-95m3/rows.tsv?accessType=DOWNLOAD&bom=true")
print(raw2)

```

# Problem 7 





```{r }
library(jsonlite)
raw3<- fromJSON("https://raw.github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/master/data/data-leg?token=AAABJG56LGWR7IB4KBYBPO3BHABSK")
print(raw3)
names<-raw3$first
surname<-raw3$last
#View(raw3)
print(names)


```


