# STAT 440 Statistical Data Management - Fall 2021
## Week03 Lab01
### Only the driver should submit this assignment into their individual student repo. Save the filename as lab01-week03-wed-netID.Rmd. Then render this file to .html when completed.

**#1** Using Markdown syntax, make a numbered list with the first and last names of everyone at your station. Your name should appear in bold text.

1. **Kamila Makhambetova**
2. Gargee Jagtap
3. Brianna Suits
4. Brianna Diaz
5. Cassidy Manuel
6. Miles Allan
7. Teresa Egan

**#2** How many lifelines are now given during the lab days?

**Answer:**: Only one lifeline is given during the lab days

**#3** Which of the following cases is the file name in bold **Homework01Kinson2**? 

a. pascal case

b. snake case

c. camel case

d. kebab case

**Answer:** Pascal case

**#4** Import the Rental Inspection Grades Listing Data using one programming language software and the data URL https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/rental-inspections-grades-data01.csv. Now, print the structure of the data. **This structure should match the data description below.** 

  - The dataset (a .csv file) contains the 1730 rental properties in Urbana with 7 variables ("columns") per property which includes details about the parcel number, address, grade, inspection date, and more. The rental properties are inspected on a rotating basis and can receive a letter grade of A, B, C, D, F, and N. The best grade is A. The original source is the City of Urbana.
  
```{r}

library(readr)
rental_inspections_grades_data01 <- read_csv("https://raw.github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/master/data/rental-inspections-grades-data01.csv?token=AAAC53B7IXLLTSVVQEQDQQTBIIHZK")
head(rental_inspections_grades_data01)
str(rental_inspections_grades_data01)

```
  

**#5** (Handling dates and times, Arranging data) Using the Rental Inspection Grades Listings Data imported in Problem **#4**, ensure that the Inspection Date and Expiration Date columns are formatted as dates or datetimes; not character format.

```{r}
library(lubridate)
colnames(rental_inspections_grades_data01) 
expdate<-rental_inspections_grades_data01$`Expiration Date`
insdate<-rental_inspections_grades_data01$`Inspection Date`
typeof(expdate) 
typeof(insdate)

rental_inspections_grades_data01$InspectionDate2<-lubridate::mdy(insdate)

rental_inspections_grades_data01$ExpirationDate2<-lubridate::mdy(expdate)

str(rental_inspections_grades_data01$InspectionDate2)
str(rental_inspections_grades_data01$ExpirationDate2)

mode(rental_inspections_grades_data01$InspectionDate2)
mode(rental_inspections_grades_data01$ExpirationDate2)
typeof(rental_inspections_grades_data01$InspectionDate2)
typeof(rental_inspections_grades_data01$ExpirationDate2)

```



**#6** Web scrape the Champaign County Assessor's [website](https://champaignil.devnetwedge.com/) to extract a property's Net Taxable Value in US Dollars. Do this for the last 10 properties in the dataset from Problem **#5** and print the resulting 10 rows.
```{r webscrapingloop}
library(rvest)
library(tidyverse)

prop_url <- paste0("https://champaignil.devnetwedge.com/parcel/view/",rental_inspections_grades_data01$`Parcel Number`,"/2020")  

owners_addresses <- rep(0, length(prop_url))

for (i in 1721:1730){ 
  owners_addresses[i] <- html_text(html_nodes(read_html(prop_url[i]),"tr:nth-child(5) td:nth-child(1) .inner-value"),trim=TRUE)
  print(owners_addresses[i])
}

```




