# STAT 440 Statistical Data Management - Fall 2021
## Week09 Lab02 Wed
### Only the driver should submit this assignment into their individual student repo. Save the filename as lab02-week09-wed-netID.Rmd (or .ipynb for Python users). Then render this file to .html when completed. 
#### Both .Rmd/.ipynb and .html files must be submitted by 11:00 am. Any lab submitted after 11:00 am will be deducted 1 point for being late.

**#1** Using Markdown syntax, make a numbered list with the first and last names of everyone at your station. Your name should appear in italic and bold text.

1. ***Kamila Makhambetova***
2. Qiuling Fan
3. Zhe Chen
4. Erh Hsuan Wang
5. Corey Shen
6. Abe Sun
7. Qixuan Xing
8. Amogh Chinnakonda

**#2** Suppose a dataset you were importing contained a date (and time) column with all entries appearing in the following format "2020-MondayNovember16". The correct format (written as percents, letters, or other characters) you should use for parsing this column is "____". **Write your answer that fills in the blank below (Markdown syntax).** *Note that the quotes are already present.*
<br>
**Answer**
<br>
format=%Y-%A%B%d


**#3** Import the following datasets and print the last 5 rows of each dataset. The datasets are part of a collection of popular baby names with the following column names: state, sex (F for female or M for male), year, name, frequency. The names are from the state of Illinois over the years 1910 to 2020. The baby names come from Social Security records. Soon after a child is born in the US, the parent(s) file for a Social Security card. This data does not include baby names with frequencies less than 5. The original source is the Social Security Administration https://www.ssa.gov/oact/babynames/limits.html.

- Iowa Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-iowa-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/yye6p5kgq0l1dtz6bqz364mr6iz3c3bw.txt).

  - The dataset (a .TXT file) contains 96176 rows and 5 columns. 

- Illinois Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-illinois-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/ttc01y8x71eadmzc9ramd16ksuq2xckk.txt).

  - The dataset (a .TXT file) contains 234613 rows and 5 columns. 

- Indiana Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-indiana-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/0ymjj08wid1l19ehy5tx9hjq2xnbse64.csv).

  - The dataset (a .csv file) contains 143083 rows and 5 columns. 

- Kentucky Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-kentucky-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/nhfxo34324uratwl3fezx08fcdoknayj.txt).

  - The dataset (a .TXT file) contains 121014 rows and 5 columns. 

- Missouri Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-missouri-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/p2synnz72z6se94int717w9jhgflmynb.csv).

  - The dataset (a .csv file) contains 142035 rows and 5 columns. 

- Wisconsin Popular Baby Names Data [GHE data URL](https://github-dev.cs.illinois.edu/stat440-fa21/stat440-fa21-course-content/raw/master/data/popular-baby-names-winconsin-data.TXT) or [Box data URL](https://uofi.box.com/shared/static/arv2amejhnyglkc684mgeyp0mcom27tl.txt).

  - The dataset (a .TXT file) contains 119521 rows and 5 columns. 
  
```{r}
library(tidyverse)
baby_names_iowa<-read_csv("https://uofi.box.com/shared/static/yye6p5kgq0l1dtz6bqz364mr6iz3c3bw.txt", col_names = FALSE)
colnames(baby_names_iowa) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_iowa, 5)

baby_names_illinois<-read_csv("https://uofi.box.com/shared/static/ttc01y8x71eadmzc9ramd16ksuq2xckk.txt", col_names = FALSE)
colnames(baby_names_illinois) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_illinois, 5)

baby_names_indiana<-read_csv("https://uofi.box.com/shared/static/0ymjj08wid1l19ehy5tx9hjq2xnbse64.csv",col_names = FALSE)
colnames(baby_names_indiana) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_indiana, 5)

baby_names_kentucky<-read_table("https://uofi.box.com/shared/static/nhfxo34324uratwl3fezx08fcdoknayj.txt",col_types="ccdcd", col_names = FALSE)
colnames(baby_names_kentucky) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_kentucky, 5)

baby_names_missouri<-read_csv("https://uofi.box.com/shared/static/p2synnz72z6se94int717w9jhgflmynb.csv",col_names = FALSE)
colnames(baby_names_missouri) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_missouri, 5)

baby_names_wisconsin<-read_csv("https://uofi.box.com/shared/static/arv2amejhnyglkc684mgeyp0mcom27tl.txt",col_names = FALSE)
colnames(baby_names_wisconsin) <- c('state','gender', 'year', 'name', 'frequency')
tail(baby_names_wisconsin, 5)


```
  

**#4** Using the 6 imported datasets in Problem **#3**, the information in the data descriptions, and a programming language, create a single dataset that contains the five columns - state, sex, year, name, and frequency - such that the rows of baby names contain no duplicates. Now, print the first 10 observations of the resulting data.

```{r}

names <- bind_rows(baby_names_illinois,baby_names_indiana,baby_names_iowa,baby_names_kentucky,baby_names_missouri,baby_names_wisconsin)
names3<-names[match(unique(names$name), names$name), ]
head(names3,10)

```


**#5** What is the average length of names (measured in characters) per state and which state had the longest average name length? Use the 6 imported datasets in Problem **#3**, the information in the data descriptions, and a programming language applying data management concepts to answer the questions. **Your answer should be in Markdown syntax and a code chunk should show the result as evidence. Duplicates names are allowed. Your result must be a single table or data frame, not multiple tables (or data frames) per state.**
<br>
<br>
**Answer**
<br>
Illinois state has the largest average length.
```{r}

mean_illinois=mean(nchar(baby_names_illinois$name))
mean_indiana=mean(nchar(baby_names_indiana$name))
mean_iowa=mean(nchar(baby_names_iowa$name))
mean_kentucky=mean(nchar(baby_names_kentucky$name))
mean_missouri=mean(nchar(baby_names_missouri$name))
mean_wisconsin=mean(nchar(baby_names_wisconsin$name))
average_length=c(mean_illinois,mean_indiana,mean_iowa, mean_kentucky,mean_missouri,mean_wisconsin)


names <- bind_rows(baby_names_illinois,baby_names_indiana,baby_names_iowa,baby_names_kentucky,baby_names_missouri,baby_names_wisconsin)
names4<-names[match(unique(names$state), names$state), ]

names4$average_length = NA

for (i in 1:length(names4)) {
  names4$average_length[i] = average_length[i]
}

names5=names4%>%select(c(state,average_length))
names5






```



